---
title: "Descriptives round 2"
author: "Monica Alexander"
date: "8/29/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Getting back into it... see what the files look like. 

```{r}
library(tidyverse)
library(lubridate)

# read in all the meta data acsv files
csv_files <- paste0("../data/raw/", list.files(path = "../data/raw",pattern = ".*.csv"))

meta_df <- c()

for(i in 1:length(csv_files)){
  this_meta <- read_csv(csv_files[i])
  meta_df <- bind_rows(meta_df, this_meta )
}
rm(this_meta)
md <- meta_df
rm(meta_df)

```

There seems to be 10,000 articles... this is less than what Leslie thought?

Format the dates and plot the number of articles by day:

```{r}
md <- md %>% 
  mutate(date = mdy(pubdate)) # thank you lubridate

md %>% 
  group_by(date) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(date, n)) + geom_line() + 
  ggtitle("Number of articles by day")
```

There's three noticable spikes: the election; the travel ban; and DACA. 

```{r}
md %>% 
  group_by(date) %>% 
  summarise(n = n()) %>% 
  arrange(-n)
```

Do a nice chart:

```{r}
md %>% 
  group_by(date) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(date, n)) + geom_line() + 
  ggtitle("Number of migration articles per day") + 
  geom_vline(xintercept = mdy("11-08-2016"), color = "red", lty = 2) +
  annotate("text", x=mdy("10-15-2016"), y=80, label= "election", color = "red", size = 5) +
  geom_vline(xintercept = mdy("01-27-2017"), color = "blue", lty = 2) +
  annotate("text", x=mdy("01-02-2017"), y=115, label= "travel ban", color = "blue", size = 5) +
  geom_vline(xintercept = mdy("09-05-2017"), color = "darkgreen", lty = 2) +
  annotate("text", x=mdy("08-17-2017"), y=55, label= "DACA", color = "darkgreen", size = 5) +
  theme_classic(base_size = 14)
ggsave("../plots/number_day.pdf", height = 7, width =10)
```

```{r, fig.width=6}
source_year <- md %>% 
  group_by(date, pubtitle) %>% 
  summarise(n = n())

source_year %>% 
  mutate(two_digit_month = ifelse(nchar(month(date))==1, paste0(0, month(date)), as.character(month(date)))) %>% 
  mutate(mon_year = paste(year(date), two_digit_month, sep = "-")) %>% 
  group_by(mon_year, pubtitle) %>% 
  summarise(n = sum(n)) %>% 
ggplot(aes(mon_year, n, color = pubtitle)) + 
  geom_point(size = 3)  + scale_color_brewer(palette="Set1", name = "Source") +
  theme_gray(base_size = 16)+
  ylab("Number") + xlab("Year-Month") + ggtitle("Number of articles by month and source") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
ggsave("../plots/source_month.pdf", height = 7, width =12)
```

Look at subject classifications. Top 20 subjects over whole period:

```{r}
# look at subjects of articles
subjects_list <- strsplit(md$subjectTerms, split = " , ")
subjects_vec <- unlist(subjects_list)
subject_counts <- sort(table(subjects_vec), decreasing = T)
head(c(subject_counts), 20)
```

Subjects by month

```{r}

md <- md %>% 
  mutate(two_digit_month = ifelse(nchar(month(date))==1, paste0(0, month(date)), as.character(month(date)))) %>% 
  mutate(mon_year = paste(year(date), two_digit_month, sep = "-")) 
dates <- unique(md$mon_year)

subject_all_years <- c()
for(i in 1:length(dates)){
  subjects_year <- unlist(subjects_list[which(md$mon_year==dates[i])])
  subjects_year_count <- sort(table(subjects_year), decreasing = T)
  subjects_year_df <- as.data.frame(subjects_year_count)
  subjects_year_df$mon_year <- dates[i]
  subject_all_years <- rbind(subject_all_years, subjects_year_df)
}

# calculate proportion 
subject_all_years <- subject_all_years %>% 
  group_by(mon_year) %>% 
  mutate(prop = Freq/sum(Freq))
```

```{r}
top5 <- subject_all_years %>% 
  group_by(mon_year) %>% 
  top_n(n= 5)

subs <- as.character(unique(top5$subjects_year))

big_subs <- subject_all_years %>% filter(subjects_year %in% subs)

ggplot(big_subs, aes(mon_year, prop, color = subjects_year)) + 
  geom_point() + geom_line() + 
  #scale_x_continuous(breaks=seq(2008,2017, by =2))+ 
  theme_minimal() + scale_color_discrete(name = "Subject") +
  ylab("Propotion") + xlab("Year") + ggtitle("Proportion of articles mentioning subject") 
```

Plot interesting subjects by day? Trump, Donald J; Aliens; Mexico; National security; Deportation; Muslims; Russia

```{r}
interesting <- c("Mexico", "National security", "Deportation", "Muslims", "Russia", "Federal court decisions", "Syria")

dates <- unique(md$date)
subject_all_years <- c()
for(i in 1:length(dates)){
  subjects_year <- unlist(subjects_list[which(md$date==dates[i])])
  subjects_year_count <- sort(table(subjects_year), decreasing = T)
  subjects_year_df <- as.data.frame(subjects_year_count)
  subjects_year_df$date <- dates[i]
  subject_all_years <- rbind(subject_all_years, subjects_year_df)
}


subject_all_years %>% 
  arrange(subjects_year,date) %>% 
  group_by(subjects_year) %>% 
  mutate(csum = cumsum(Freq)) %>% 
  arrange(subjects_year,csum) %>% 
  filter(subjects_year %in% interesting) %>% 
  ggplot(aes(date, csum, color = subjects_year)) + geom_line(lwd = 1.2)+
  ggtitle("Cumulative number of articles covering selected subjects")+
  scale_color_discrete(name = "Subject") +
  ylab("Cumulative total") + 
  theme_classic(base_size = 14)
ggsave("../plots/cumulative_topics.pdf", height = 7, width =10)
 ```

